---
title: "Homework 12"
author: "Peyton Hall"
date: "04/11/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Load necessary libraries}
library(stats)
library(readxl)
```


1. A clinic developed a new scheduling method to reduce the amount of unscheduled nurse visits.
The proportion of unscheduled nurse visits was 40% before this intervention. After this
intervention, in a sample of 100 clinic visits, 25 of them were unscheduled nurse visits. Use the
0.05 significance level to test the claim that the proportion of unscheduled nurse visits is
significantly less than 40%. Show your R code. Write down the H0 and Ha and the decision to
address the original claim in the context in R Markdown using comments.
$H_0: P=.40$ vs $H_A: P<.40$
```{r Question 01}
#(x,n,p)
prop.test(25,100,p=.4,alternative = "less")
```
The p value is 0.001539. The P-Value is low because, compared to the 
significance level, it is less than .05. Therefore, we reject $H_0$
and go with the $H_A$.
There is sufficient evidence to support the claim that the proportion of 
unscheduled nurse visits is significantly less than 40%.


2. A pharmaceutical company developed a new drug (A) to treat migraine. The company would like
to compare the proportion of people who responded positively to drug A is significantly higher
than the proportion of people who responded positively to the standard drug (drug B). On one
sample of 200 people who took A, 130 responded positively, and on another sample of 100 people
who took B, 50 people responded positively. Use the 0.05 significance level to test the claim that
the proportion of people who responded positively to A is significantly higher than the proportion
who responded positively to B. Show your R code. Write down the H0 and Ha and the decision to
address the original claim in the context in R Markdown using comments.
$H_0: P_A = P_B$ vs $H_A: P_A > P_B$
let p_A = Drug A
let p_B = Drug B
```{r Question 02}
AB <- c(130, 50)
N_AB <- c(200, 100)
prop.test(AB, N_AB, alternative = "greater")
```
The probability value is 0.008774. That is, lower than the 0.05 significance 
level. Therefore, we reject the $H_0$ and go with the $H_A$.
There is sufficient evidence to support the claim that the proportion of people 
who responded positively to A is significantly higher than the proportion
who responded positively to B


3. A car manufacturer aims to improve the quality of the products by reducing defects and hence
increasing customer satisfaction. He monitors the efficiency of two assembly lines in the shop. In
line A, there are 18 defects reported out of 200 samples. In line B, there are 25 defects out of 600
cars. Use the 0.05 significance level to test the claim that the proportion of defects in line A is
significantly different from the proportion in line B. Write down the H0 and Ha, and the decision
to address the original claim in the context in R Markdown using comments.
$H_0: P_A = P_B$ vs $H_A: P_A \ne P_B$
```{r Question 03}
car_AB <- c(18, 25)
car_N_AB <- c(200, 600)
prop.test(car_AB, car_N_AB, alternative = "two.sided") # syntax for not equal
```
The p-value is 0.01453. The significance level is 0.05. The p-value is less than
the significance level. Therefore, we would reject the $H_0$ and go with the 
$H_A$. There is sufficient evidence to support the claim that the proportion of 
defects in line A is significantly different from the proportion in line B.


4. The data “COVID19 Vaccine Preference” on D2L showed a sample of 50 individuals’ preference
for the type of COVID19 vaccine. Use the table() function to get the counts of each preference
first. Use the 0.05 significance level to test the claim that the proportion of preferring Moderna
vaccine is significantly different from the proportion of preferring Pfizer. (use a comment to show
your null and alternative hypothesis and the decision to the claim)
$H_0: P_M = P_Z$ vs $H_A: P_M \ne P_Z$
let P_M = proportion of Moderna preference
let P_Z = proportion of Pfizer preference
```{r Question 04}
COVID19_Vaccine_Preference <- read_excel("~/Desktop/Data211/Week 13/COVID19 Vaccine Preference.xlsx")
COVID19_Vaccine_Preference
table(COVID19_Vaccine_Preference$Preference)
length(COVID19_Vaccine_Preference$Preference)

Preference <- c(26,24) # total of 50
V <- c(50,50) 

prop.test(Preference,V,alternative = "two.sided") #(x,n,p)
```
The p-value is 0.8415. The significance level is 0.05. The p-value is greater
than the significance level. Therefore, we fail to reject the $H_0$.
There is insufficient evidence to support the claim that the proportion of 
preferring Moderna vaccine is significantly different from the proportion of 
preferring Pfizer.


5. The data “onlineLearning” on D2L recorded the students’ ratings of synchronous online classes
and asynchronous online classes. Use the 0.05 significance level to test the claim that the
proportion of a rating of 5 for synchronous classes is significantly higher than the proportion of a
rating of 5 for asynchronous classes. Use table() to get the corresponding counts.
$H_0: P_S = P_A$ vs $H_A: P_S > P_A$
```{r Question 05}
onlineLearning <- read_excel("~/Desktop/Data211/Week 13/onlineLearning.xlsx")
onlineLearning
n<-as.numeric(length(onlineLearning$ID))
n<-c(n,n)
x<-c(as.numeric(table(onlineLearning$Async)[5]), as.numeric (table(onlineLearning$Sync)[3]))
prop.test (x, n, alternative = "less")
```
The p-value is 0.0153 The significance level is 0.05. The p-value is 
significantly lower than the significance level. Therefore, we would reject the
$H_0$ and go with the $H_A$.
There is sufficient evidence to support the claim that the
proportion of a rating of 5 for synchronous classes is significantly higher than 
the proportion of a rating of 5 for asynchronous classes.
