---
title: "Week13InClassLecture"
author: "Peyton Hall"
date: "04/04/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load necessary libraries}
library(stats)
library(readxl)
```

$H_0:P = 0.2$ vs $H_a: P > 0.2$
```{r}
prop.test(310,1000,p=0.2,alternative = "greater")
```
the p-value < 2.2e-16, which is extremely low. Therefore, we should reject the 
$H_0$ and go with $H_a$.

$H_0: P = 0.5$ vs $Ha: P > 0.5$
write doen the null and alternative. after the R chunk, write down the conclusion.
```{r}
prop.test(8,8,0.5,alternative = "greater")
```
the p-value = 0.006664. If the significance level is < 0.001, then we reject the p-value.

$H_0:P=0.75$ vs $H_a:P>0.75$
```{r Class Poll Activity}
# how manuy people can roll their tongues?
prop.test(14,16,0.75,alternative = "greater")
```
p-value=0.1932>0.05
We fail to reject the $H_0$.
At 5% significance level data do not provide
enough evidence to conclude proportion of students
who can roll their tongue is greater than 75%

p:Proportion of drivers who feels comfortable driving self driving car
$H_0:P=0.5$ vs $H_a:P<0.5$
```{r Slide 27 Activity 02}
selfDriving <- read_excel("~/Desktop/Data211/Week 13/Responses_to_Self-driving_Cars.xlsx")
selfDriving
table(selfDriving$Comfortable)

length(selfDriving$Comfortable)#this will include NAs

prop.test(2860,5962,0.5,alternative = "less")
```
p-value = 0.000906 < 0.01
We reject $H_0$ at 1% significance level
At 1% s.l, we have enough evidence to conclude that the P<50%

```{r}
prop.test(x=c(51,19),n=c(303,303),alternative = "greater")
```
p=4.079e-05<0.01

Reject the $H_0$.
At 1% s.l we have enough evidence to conclude that the proportion of voters
who favor in YS is significantly higher than those who favor on Yo.

