---
title: "Ulcerative Colitis Weight Rehabilitation"
author: "Peyton Hall"
date: "04/25/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Necessary Libraries}
library(ggplot2)
```

```{r Create Columns & Rows}
# Create a matrix UC
UC <- matrix(NA, nrow = 14, ncol = 4)

# Assign the data to the matrix
UC <- matrix(
  c("02/24/2024", "148.6", "4 Pills (40MG)", NA,
    "02/25/2024", "150.2", "4 Pills (40MG)", NA,
    "02/26/2024", "151.6", "4 Pills (40MG)", "Visited Gym",
    "02/27/2024", "152.4", "4 Pills (40MG)", NA,
    "02/28/2024", "154.2", "4 Pills (40MG)", NA,
    "02/29/2024", "156.6", "4 Pills (40MG)", NA,
    "03/01/2024", "156.4", "4 Pills (40MG)", "Visited Gym",
    "03/02/2024", "158.0", "3 Pills (30MG)", NA,
    "03/03/2024", "158.2", "3 Pills (30MG)", NA,
    "03/04/2024", "158.8", "3 Pills (30MG)", "Visited Gym",
    "03/05/2024", "158.8", "3 Pills (30MG)", NA,
    "03/06/2024", "157.4", "3 Pills (30MG)", NA,
    "03/07/2024", "160.6", "3 Pills (30MG)", "Visited Gym",
    "03/08/2024", "158.4", "3 Pills (30MG)", "Visited Gym"),
  nrow = 14,
  byrow = TRUE
)

# Assign column names
colnames(UC) <- c("Date", "Weight", "Prednisone Dosage", "Activity")

# Print the matrix
UC
```

## Note: Prednisone Chemical Formular = C_21 H_26 O_5

```{r Visualize the Data}
# Convert matrix UC to data frame
UC_df <- as.data.frame(UC)

# Convert weight column to numeric, handling non-numeric entries
UC_df$Weight <- as.numeric(ifelse(grepl("[0-9.]+", UC_df$Weight), 
                                  UC_df$Weight, NA))

# Create the scatter plot with regression line using the data frame UC_df
ggplot(UC_df, aes(x = as.Date(Date, "%m/%d/%Y"), y = Weight, 
                  color = as.factor(`Prednisone Dosage`))) +
  geom_point() +  # Scatter plot points
  guides(color = guide_legend(title = "Prednisone Dosage")) +
  scale_color_manual(values=c("#01BAB3", "#BF2020")) +
  geom_smooth(method = "lm", se = FALSE, color = "#005587") +  # Regression line
  labs(title = "Ulcerative Colitis Weight Rehabilitation", 
       x = "Date (m/d) 2024", y = "Weight (lbs)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +  # Center the title
  scale_x_date(date_breaks = "1 day", date_labels = "%m/%d")
```

## Hypothesis: The average weight gain rate was higher on 4 pills rather than 3.
$H_0: \mu f = \mu t$ vs $H_a: \mu f > \mu t$
Let f = average rate of change of weight on four pills
Let t = average rate of change of weight on three pills
```{r}
UC_df$days <- c(1:14) # create a vector 1 through 14 and save it into the UC_df data frame

model.4days = lm(Weight ~ days, data = UC_df[1:7,])
model.3days = lm(Weight ~ days, data = UC_df[8:14,])

summary(model.4days)
summary(model.3days)

# average rate of change formula = (y2 - y1) / (x2 - x1)
# average rate of change = (weight2 - weight 1) / (time2 - time1)
f = c(1.6, 1.4, 0.8, 1.8, 2.4, -0.2)
t = c(0.2, 0.6, 0.0, -1.4, 3.2, -2.2)

# use paired t-test because there is a difference in medication per week
t.test(f, t, alternative = "greater", paired = TRUE) # (x,n,p)

```
The probability value is 0.03702. The significance level can be defined as 
alpha = 0.05. The mean rate of change for the first set is greater than the mean 
rate of change for the second set. That is, the mean difference equals 1.233333.
The $H_0$ should be rejected, and the $H_a$ should be accepted. 
There is significant evidence that the average rate of change for four pills is 
greater than the average rate of change for three pills.
